\documentclass[8pt, twocolumn]{article}
\usepackage{hyperref}
\usepackage{times}
\hypersetup{
colorlinks=true,
linkcolor=blue,
citecolor=blue,
urlcolor=blue,
}

\title{Concurrent Programming in Go \\
\small{\url{http://www.golang.org}}
} 
\author{
  Raghu Prabhakar
  \\\small{raghu@cs.ucla.edu}
  \and 
  Rohit Kumar 
  \\\small{rohitkk@cs.ucla.edu}
}

\begin{document}

\maketitle
\section{Abstract}
The Go programming language was released by Google in November
2009. The language was designed and implemented by the creators of
Unix: Rob Pike and Ken Thompson. It's a general purpose programming
language intended for systems programming. This manual gives a brief
overview of the language's support for concurrency.\\
\section{Introduction}
Go was designed as a general purpose system's programming language. It
borrows from C (pointers, goto), Newsqueak (channels) \cite{gofaq}. It
has garbage collection, is statically typed and has compilers on
Windows, Linux and MacOS X.\\

The primitives for concurrency present in Go are similar to those
in CSP \cite{goblog}. Goroutines are used to spawn new threads of
execution and channels are used to communicate between
goroutines. Although this model of concurrency is the recommended one,
the programmer can choose to use shared memory and locks. The
following sections talk about Go's concurrency primitives and their
usage.

\section{Goroutines}
Goroutines are functions executing in a separate thread. To start a
goroutine, you prefix the function call with the \verb=go= keyword.
\begin{verbatim}
        go map_block(start, end)
\end{verbatim}
The above statement will start the \verb=map_block= function as a
goroutine in a separate thread. This is an asynchronous call and the
control will not wait for the execution of \verb=map_block= to finish
before executing the next statement\cite{golangspec}. When the
goroutine finishes it will exit silently. The goroutine shares the
same memory as other goroutines and the main thread of
execution. Multiple goroutines might be multiplexed on the same system
thread.\\

By default the Go runtime will only use one processor to schedule
goroutines. To use more than one processor, one should call the
\verb=runtime.GOMAXPROCS= function. For example:
\begin{verbatim}
        import ("runtime")        
        func main() {
                runtime.GOMAXPROCS(4)
        }
\end{verbatim}
The above code instructs the runtime to use 4 processors.

\section{Channels}
Channels are the main form of syncrhonization provided by Go. They can
be used to send and receive values between goroutines. Channels are
typed.
\begin{verbatim}
    1    ch := make(chan int)                         
    2    go func() {
    3            v:= <-ch)
    4    }()
    5    ch <- 23  
\end{verbatim}
On Line 1 we create a new channel using \verb=make=. The default
channels are unbuffered and will block on both send and receive. We
then spawn a new goroutine which receives a value from the channel
(Line 3). Finally we send a the number \verb=23= through the channel
(Line 5).\\

To send a value through a channel one uses the \verb=<-= operator with the
channel on the left hand side (Line 5). To receive a value, place the
channel on the right hand side of the \verb=<-= operator. \\

The order of send and receive is important. If we had done
\verb=ch<-23= before Line 2. The program would block and never execute
the \verb=go= statement. As an unbuffered channel blocks
both on a send and a receive \cite{effectivego}.

Channels can also be used to wait for goroutines to finish. For
example:
\begin{verbatim}
    1     ch := make(chan bool)
    2     for i:=0; i<n; i++ {
    3             go func() {
    4             //do something
    5                     ch <- true
    6             }()
    7     }
    8     for i:=0; i<n; i++ {
    9             <-ch 
    10    } 
\end{verbatim}
In the above code snippet we spawn an anonymous goroutine \verb=n=
times (Line 3). Each one of these goroutines sends a value \verb=true=
on the channel \verb=ch= after it has finished its task (Line 5). In
the main goroutine we wait for all child goroutines to finish by
receiving the values on the channels and discarding them (Line 9). We
know that after Line 10, all the spawned goroutines have finished.\\

There are other patterns of using channels, many of
which are illustrated in \cite{effectivego}.
\section{WaitGroup}
Waitgroups are a better construct to syncrhonize the completion of
goroutines. They are present in the \verb=sync= package. The same code
listed in the previous section can be rewritten using a
\verb=WaitGroup=. 
\begin{verbatim}
    1     var wg sync.WaitGroup
    2     for i:=0; i<n; i++ {
    3             wg.Add(1)
    4             go func() {
    5             //do something
    6                     wg.Done()
    7             }()
    8     }
    9     wg.Wait()
    10    } 
\end{verbatim}
Here we see that the main goroutine calls add to set the number of
goroutines to wait for (Line 3). When each goroutine has finished
execution, it calls the \verb=Done= method (Line 6) on the
\verb=WaitGroup=. The main routine waits for the all the child
goroutines to finish by calling \verb=Wait= (Line 9) \cite{sync}.

\section{Select}
The \verb=select= statement is used to choose a send or receive from a
set of channels. The statement is structured like a \verb=switch=
statement, with each case being either a send to or a receive from a
channel. Each of these cases is evaluated from the top to bottom. Out
of all the send/receive expressions which can proceed one is selected
and executed. See \cite{golangspec} for examples.
\section{Locks}
Within the \verb=sync= package, which comes with the standard Go
distribution, there are two types of locks: Mutex and Reader Writer
Lock. The recommended way of synchronizing is channels. The locks
provided here are used to build higher level synchronization
mechanisms. See \cite{sync} for more details.
\section{Once}
The \verb=Once= structure can be used to execute a particular function
only one time. For example:
\begin{verbatim}
    1     var once sync.Once
    2     for i:=0; i<n; i++ {
    3             go func() {
    4                     //do something
    5                     once.Do(cleanup)
    6             }()
    9     }
\end{verbatim}
In the above code although multiple goroutines will reach Line 5, only
one of them will get to execute the \verb=cleanup= function.
\section{Conclusion}
The Go programming language has been around for almost 2 years. Although
it was experimental when it was released it has since been deployed in
production environments \cite{heroku}. Given that it's general
purpose, imperative, easy to learn and provides native concurrency
primitives, we believe that it will be adopted by many more developers in the
future.
\nocite{*}
\bibliography{manual} 
\bibliographystyle{plain}
\end{document}
